Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-08-07 alloscore_manuscript/manuscript.Rnw
bisection explanation

\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations such that $F_1$ and $F_2$ have exponential marginal distribution with scale parameters $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
$\sigma_i$ (see appendix \ref{sec:a:bayes-quantiles}). If $K = 5$ units of the resource are
available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
Increasing $K$ to $10$ changes this optimal allocation to 2 and 8. Figure~\ref{fig:exp_alloc_example} illustrates the 
situation.

@@

\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in
each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale
parameters of the forecast in Example 1. Because the optimal allocation is proportional to the scale parameters, this
forecast would lead to the same allocations as the forecast in Example 1, and would therefore be assigned the same
allocation score.

@@

Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:AP} in section
\ref{sec:methods.detailed.specific_allocation}) of the text:
\begin{align}
    \underset{0 \leq x}{\mathrm{minimize}}\,\, 
    \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]
     \text{ subject to }
     \, \sum_{i=1}^N x_i = K, \label{eqn:AP-long}
\end{align}
where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations. We show that the
Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$ is a vector of
quantiles of the marginal forecast distributions $F_i$ at a single probability level $\tau^{F,K}$, that is, $x_i^{F,K} =
q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref {sec:methods.overview} is that if
$F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is proportional to $(\sigma_1,\ldots,\sigma_N)$, since
$q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.

@@

To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a Bayes act solving the
AP \eqref{eqn:AP-long} --- that is, we must find the specific resource allocations for each location that are determined by
the forecast $F$ under the resource constraint $K$. Assuming we have reliable means of calculating quantiles
$q_{\alpha,F_i}$ of the marginal forecasts $F_i$, these allocations are given by $q_{\tau^{\star},F_i}$ where
$\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}. However, this equation is not analytically tractable
and we must resort to a numerical method for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.

We have implemented an iterative bisection method that makes use of the fact that $\sum_{i=1}^N q_{\tau,F_i}$ is an
increasing function of $\tau$. The algorithm begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as
$[0,\max_{i}F_i(K)]$) that clearly contains the solution $\tau^{\star}$. At each step $j$ of the algorithm, we evaluate
the total allocation $\sum_{i=1}^N q_{\tau_{M,j},F_i}$ at the midpoint of the search interval,
$\tau_{M,j} = \frac{1}{2}(\tau_{L,j} + \tau_{U,j})$ and continue the search on the narrowed sub-interval
\begin{align}
[\tau_{L,j+1},\tau_{U,j+1}] =
\begin{cases}
[\tau_{L,j}, \tau_{M,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} \geq K \\
[\tau_{M,j}, \tau_{U,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} < K.
\end{cases} \nonumber
\end{align}
This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ for a suitably small $\varepsilon>0$. We have
implemented this procedure along with the resulting score computations in the R package \verb`alloscore`
\citep{gerding-alloscore} which provided all allocation score values used in the analysis of section
\ref{sec:application}.

>>>>

\paragraph{Example 1} \label{examp1} Suppose we have a forecast $F$ for need in two locations such that $F_1$ and $F_2$ have exponential marginal distribution with scale parameters $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
$\sigma_i$ (see appendix \ref{sec:a:bayes-quantiles}). If $K = 5$ units of the resource are
available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
Increasing $K$ to $10$ changes this optimal allocation to 2 and 8. Figure~\ref{fig:exp_alloc_example} illustrates the 
situation.

@@

\paragraph{Example 2} \label{examp2} Now consider a different forecast that also has exponential distributions for resource need in
each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale
parameters of the forecast in Example 1. Because the optimal allocation is proportional to the scale parameters, this
forecast would lead to the same allocations as the forecast in Example 1, and would therefore be assigned the same
allocation score.

@@

Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:AP} in section
\ref{sec:methods.detailed.specific_allocation}) of the text:
\begin{align}
    \underset{0 \leq x}{\mathrm{minimize}}\,\, 
    \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]
     \text{ subject to }
     \, \sum_{i=1}^N x_i = K, \label{eqn:AP-long}
\end{align}
where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations. We show that the
Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$ is a vector of
quantiles of the marginal forecast distributions $F_i$ at a single probability level $\tau^{F,K}$, that is, $x_i^{F,K} =
q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref {sec:methods.overview} is that if
$F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is proportional to $(\sigma_1,\ldots,\sigma_N)$, since
\begin{align}
q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau). \label{eqn:exp-quantile}
\end{align}

@@

To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a Bayes act solving the
AP \eqref{eqn:AP-long} --- that is, we must find the specific resource allocations for each location that are determined by
the forecast $F$ under the resource constraint $K$. Assuming we have reliable means of calculating quantiles
$q_{\alpha,F_i}$ of the marginal forecasts $F_i$, these allocations are given by $q_{\tau^{\star},F_i}$ where
$\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}. This equation is not analytically tractable in general, but because the left-hand side of \eqref{eqn:quantiles-sum-to-K} is non-decreasing in $\tau$, it is straightforward to find arbitrarily accurate approximations to a solution using an iterative bisection method.

We have
implemented such a method along with the resulting score computations in the R package \verb`alloscore`
\citep{gerding-alloscore} which provided all allocation score values used in the analysis of section
\ref{sec:application}.
The procedure begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as
$[0,\max_{i}F_i(K)]$) that clearly contains the solution $\tau^{\star}$. At each step $j$ of the algorithm, we evaluate
the total allocation $\sum_{i=1}^N q_{\tau_{M,j},F_i}$ at the midpoint of the search interval,
$\tau_{M,j} = \frac{1}{2}(\tau_{L,j} + \tau_{U,j})$ and continue the search on the narrowed sub-interval
\begin{align}
[\tau_{L,j+1},\tau_{U,j+1}] =
\begin{cases}
[\tau_{L,j}, \tau_{M,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} \geq K \\
[\tau_{M,j}, \tau_{U,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} < K.
\end{cases} \nonumber
\end{align}
This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ 
for a suitably small $\varepsilon>0$.

Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref {sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f. \eqref{eqn:exp-quantile}).

<<bisection, fig.height=4, fig.width = 8, fig.pos="H", fig.cap = "Illustrations of bisection method for the examples of Section \\ref{sec:methods.overview}. In each plot the thinner black lines are the quantile functions of the exponential forecasts for Location 1, with $\\sigma_1=1$ (lower), and Location 2, with $\\sigma_2=4$ (upper), while the thicker black line gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right. The red segments show the iterates $\\tau_{M,j}$ and their associated total resource demands as given by the function \\texttt{allocate} from the package \\texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$. (The $\\tau_{M,j}$ are in this case just sums of terms $\\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f. Figure \\ref{fig:exp_alloc_example} ).">>=
expexamp <- tibble(
  target_names = LETTERS[1:2],
  dist = "exp",
  rate = c(1,.25),
) %>% add_pdqr_funs(types = c("p", "d", "q"))

alloc <- allocate(df = expexamp, K = c(5,10))

q1 <- expexamp$Q[[1]]
q2 <- expexamp$Q[[2]]
q_sum <- function(tau) q1(tau) + q2(tau)
iters5 <- tibble(taus = 1-alloc$lam_seq[[1]], qtot = q_sum(taus))
iters10 <- tibble(taus = 1-alloc$lam_seq[[2]], qtot = q_sum(taus))

p <- ggplot() +
  geom_function(fun = q1, n=1000) +
  geom_function(fun = q2, n=1000) +
  geom_function(fun = q_sum, n=1000, linewidth = 1) +
  labs(
    x = expression(tau),
    y = expression(paste(italic(q[tau * "," * F[1]]),
                         ", ", 
                         italic(q[tau * "," * F[2]]),
                         "  and  ", 
                         italic(q[tau * "," * F[1]] + q[tau * "," * F[2]])))) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.length.x = unit(.25, "cm"))

p5 <- p + geom_hline(aes(yintercept = 5)) +
  geom_hline(aes(yintercept = c(1,4)), linetype = 2) +
  geom_segment(data = iters5, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
                     breaks = c(.5,.6,0.632,.7,.8,.9,1),
                     labels = c("0.5","0.6",expression(italic(atop(" ", 1 - e^{-1} %~~% 0.632))),
                                "0.7","0.8","0.9","1")) +
  scale_y_continuous(limits = c(0,12), breaks = c(0,1,4,5,10,15), expand = c(0,0))

p10 <- p + geom_hline(aes(yintercept = 10)) +
  geom_hline(aes(yintercept = c(2,8)), linetype = 2) +
  geom_segment(data = iters10, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
                     breaks = c(.5,.6,.7,.8,0.865,.9,1),
                     labels = c("0.5","0.6",
                                "0.7","0.8", 
                                expression(italic(atop(" ", 1 - e^{-1} %~~% 0.865)), "0.9", "1"))) +
  scale_y_continuous(limits = c(0,12), breaks = c(0,2,5,8,10,15), expand = c(0,0), 
                     name = element_blank())

cowplot::plot_grid(p5, p10, align = "h")

### 2024-08-07 alloscore_manuscript/manuscript.tex
bisection explanation

The allocation score $S_A$ developed in the previous section evaluates the forecast distributions
$F$ based on a single probability level $\tau^{F,K}$. This is appropriate if the resource constraint $K$ is a known
constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the
value of forecasts across a range of decision making scenarios with different resource constraints, we can use an
\emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a
distribution $p$:
\begin{align} 
S_{IAS}(F, y) &= \int S_A(F,y; K) p(K) \, dK \label{eq:IASintegral}\\
	& = \sum_{K=1}^{\infty} S_A(F,y; K) p(K) \label{eq:IASsum}.
\end{align}
We note that the device of considering a range of hypothetical decision makers or decision making problems with
different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.

@@

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Thu Aug  1 13:13:10 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
\texttt{CU-select} & 669 & 133 & 774 & 326 \\ 
  \texttt{per-capita} & 865 & - & 1029 & 366 \\ 
  \texttt{COVIDhub-ensemble} & 873 & 159 & 1067 & 438 \\ 
  \texttt{USC-SI\_kJalpha} & 995 & 91 & 1216 & 1097 \\ 
  \texttt{JHUAPL-Gecko} & 1034 & 164 & 1141 & 418 \\ 
  \texttt{MUNI-ARIMA} & 1084 & 169 & 1248 & 440 \\ 
  \texttt{COVIDhub-trained\_ensemble} & 1089 & 169 & 1271 & 823 \\ 
  \texttt{COVIDhub-baseline} & 1175 & 170 & 1317 & 535 \\ 
  \texttt{JHUAPL-Bucky} & 1358 & 102 & 1566 & 1214 \\ 
  \texttt{JHUAPL-SLPHospEns} & 1540 & 129 & 1604 & 1102 \\ 
  \texttt{UVA-Ensemble} & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), (see Section \ref{sec:ias_examp}). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

@@

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Thu Aug  1 13:13:14 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
model & MAS & taMWIS & taMWIS rank \\ 
  \hline
\texttt{COVIDhub-ensemble} & 389 & 70 & 2 \\ 
  \texttt{per-capita} & 464 & - & - \\ 
  \texttt{COVIDhub-trained\_ensemble} & 483 & 87 & 4 \\ 
  \texttt{CU-select} & 502 & 81 & 3 \\ 
  \texttt{JHUAPL-SLPHospEns} & 526 & 67 & 1 \\ 
  \texttt{COVIDhub-baseline} & 594 & 93 & 5 \\ 
  \texttt{JHUAPL-Bucky} & 643 & 112 & 7 \\ 
  \texttt{MUNI-ARIMA} & 707 & 116 & 8 \\ 
  \texttt{JHUAPL-Gecko} & 929 & 110 & 6 \\ 
  \texttt{USC-SI\_kJalpha} & 1473 & 155 & 9 \\ 
   \hline
\end{tabular}
\caption{Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. MWIS is not defined for the \texttt{per-capita} allocation which does not use forecasts.} 
\label{tab:multi-week-performance-summary}
\end{table}

@@

To assess the sensitivity of AS to choice of $K$, AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A). These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of 19,581 that day. Model ranks by AS were fairly stable across all $K$, and there appears to be an substantial interval around the observed value over which ranks are constant.

@@

The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
constraint ($K$), allowing one to assign higher weights to more likely values of $K$ (Section
\ref{sec:methods.detailed.integrated_allocation}). Calculating exact IAS as defined in equation \eqref{eq:IASsum} is computationally intensive so an approximate IAS, denoted $\hat{S}(F,y)$, was used such that
\begin{align*}
	\hat{S}(F,y) = \sum_{k=1}^{600} S_A(F,y; 100k) p(100k),
\end{align*}
where summation terms were calculated at every 100 beds with an upper bound of 60,000, or four times the estimated available beds.

Approximate IAS was computed for two weight distributions on $K$, one uniform
across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
\ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
the higher correlation coming from the weighting centered at $K=15,000$ (Figure \ref{fig:multi-k}B). Model rankings
based on the AS and the centered IAS were roughly similar, with the top and bottom three approaches being the same for
both scores (Table \ref{tab:multi-k-scores}).

@@

Additionally, during an epidemic, public health decision makers may be faced with an ongoing series of decision over time in response to changes in conditions such as resource availability or the current allocations of resources. The AS presented here looks at a single snapshot in time  and ignores many of the complexities related to adjusting existing allocations. The example presented in section \ref{sec:application} is based on a prospective estimate of bed availability, but one could also use observed total resource availability to derive an optimal set of prospective AS parameters. In the case of allocation of preventative resources, such as vaccines, one may which to account for the negative feedback loop between resource allocation and disease cases. Alternatively, the impact of unmet need often differs across different locations and demographics, which could be accounted for by incorporating a function for the cost of unmet need, which differs by location or other covariates. Future work is needed to explore the use of AS while accounting for the dynamic nature of disease spread and resource availability.

@@

In conclusion, we argue that when possible, the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
the specific decision-making context. Defaulting to standard forecast evaluation metrics can mask the utility (or
disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
for which they are able to offer useful guidance. The work presented here New collaborative work between decision-makers and modeling teams is
needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or
simulation exercises that could be used to inform further development of new or alternative scoring metrics. We see this
work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling
applied epidemiological forecasting with public health decision making.

@@

To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a Bayes act solving the
AP \eqref{eqn:AP-long} --- that is, we must find the specific resource allocations for each location that are determined by
the forecast $F$ under the resource constraint $K$. Assuming we have reliable means of calculating quantiles
$q_{\alpha,F_i}$ of the marginal forecasts $F_i$, these allocations are given by $q_{\tau^{\star},F_i}$ where
$\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}. However, this equation is not analytically tractable
and we must resort to a numerical method for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.

We have implemented an iterative bisection method that makes use of the fact that $\sum_{i=1}^N q_{\tau,F_i}$ is an
increasing function of $\tau$. The algorithm begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as
$[0,\max_{i}F_i(K)]$) that clearly contains the solution $\tau^{\star}$. At each step $j$ of the algorithm, we evaluate
the total allocation $\sum_{i=1}^N q_{\tau_{M,j},F_i}$ at the midpoint of the search interval,
$\tau_{M,j} = \frac{1}{2}(\tau_{L,j} + \tau_{U,j})$ and continue the search on the narrowed sub-interval
\begin{align}
[\tau_{L,j+1},\tau_{U,j+1}] =
\begin{cases}
[\tau_{L,j}, \tau_{M,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} \geq K \\
[\tau_{M,j}, \tau_{U,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} < K.
\end{cases} \nonumber
\end{align}
This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ for a suitably small $\varepsilon>0$. We have
implemented this procedure along with the resulting score computations in the R package \verb`alloscore`
\citep{gerding-alloscore} which provided all allocation score values used in the analysis of section
\ref{sec:application}.

>>>>

The allocation score $S_A$ developed in the previous section evaluates the forecast distributions
$F$ based on a single probability level $\tau^{F,K}$. This is appropriate if the resource constraint $K$ is a known
constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the
value of forecasts across a range of decision making scenarios with different resource constraints, we can use an
\emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a
distribution $p$:
\begin{align} 
  S_{IAS}(F, y) &= \int S_A(F,y; K) p(K) \, dK \label{eq:IASintegral}\\
    & = \sum_{K=1}^{\infty} S_A(F,y; K) p(K) \quad (\text{assumming non-divisible resource units})\label{eq:IASsum}.
\end{align}
We note that the device of considering a range of hypothetical decision makers or decision making problems with
different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.

@@

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Aug  7 14:47:41 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
\texttt{CU-select} & 669 & 133 & 774 & 326 \\ 
  \texttt{per-capita} & 865 & - & 1029 & 366 \\ 
  \texttt{COVIDhub-ensemble} & 873 & 159 & 1067 & 438 \\ 
  \texttt{USC-SI\_kJalpha} & 995 & 91 & 1216 & 1097 \\ 
  \texttt{JHUAPL-Gecko} & 1034 & 164 & 1141 & 418 \\ 
  \texttt{MUNI-ARIMA} & 1084 & 169 & 1248 & 440 \\ 
  \texttt{COVIDhub-trained\_ensemble} & 1089 & 169 & 1271 & 823 \\ 
  \texttt{COVIDhub-baseline} & 1175 & 170 & 1317 & 535 \\ 
  \texttt{JHUAPL-Bucky} & 1358 & 102 & 1566 & 1214 \\ 
  \texttt{JHUAPL-SLPHospEns} & 1540 & 129 & 1604 & 1102 \\ 
  \texttt{UVA-Ensemble} & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), (see Section \ref{sec:ias_examp}). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

@@

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Aug  7 14:47:46 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
model & MAS & taMWIS & taMWIS rank \\ 
  \hline
\texttt{COVIDhub-ensemble} & 389 & 70 & 2 \\ 
  \texttt{per-capita} & 464 & - & - \\ 
  \texttt{COVIDhub-trained\_ensemble} & 483 & 87 & 4 \\ 
  \texttt{CU-select} & 502 & 81 & 3 \\ 
  \texttt{JHUAPL-SLPHospEns} & 526 & 67 & 1 \\ 
  \texttt{COVIDhub-baseline} & 594 & 93 & 5 \\ 
  \texttt{JHUAPL-Bucky} & 643 & 112 & 7 \\ 
  \texttt{MUNI-ARIMA} & 707 & 116 & 8 \\ 
  \texttt{JHUAPL-Gecko} & 929 & 110 & 6 \\ 
  \texttt{USC-SI\_kJalpha} & 1473 & 155 & 9 \\ 
   \hline
\end{tabular}
\caption{Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. MWIS is not defined for the \texttt{per-capita} allocation which does not use forecasts.} 
\label{tab:multi-week-performance-summary}
\end{table}

@@

AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A). These
calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
19,581 that day. Model ranks by AS were fairly stable across all $K$, and there appears
to be an substantial interval around the observed value over which ranks are constant.

@@

The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
constraint ($K$), allowing one to assign higher weights to more likely values of $K$ (Section
\ref{sec:methods.detailed.integrated_allocation}). 
Calculating exact IAS as defined in equation \eqref{eq:IASsum} is computationally intensive so an approximate IAS, denoted $\hat{S}(F,y)$, was used such that
\begin{align*}
	\hat{S}(F,y) = \sum_{k=1}^{600} S_A(F,y; 200k) p(200k),
\end{align*}
where summation terms were calculated at every 200 beds with an upper bound of 60,000, or four times the estimated available beds.
An approximate IAS was computed for two weight distributions on $K$, one uniform
across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
\ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
the higher correlation coming from the weighting centered at $K=15,000$ (Figure \ref{fig:multi-k}B). Model rankings
based on the AS and the centered IAS were roughly similar, with the top and bottom three approaches being the same for
both scores (Table \ref{tab:multi-k-scores}).

@@

Additionally, during an epidemic, public health decision makers may be faced with an ongoing series of decisions over time in response to changes in conditions such as resource availability or the current effective allocations of resources. The AS presented here looks at a single snapshot in time  and ignores many of the possible complexities related to adjusting existing allocations. The example presented in section 
\ref{sec:application} is based on a prospective estimate of bed availability, but one could also use observed total resource availability to target an optimal set of prospective AS parameters. 
Alternatively, the impact of unmet need often differs across different locations and demographics, which could be accounted for by incorporating a function of location and demographics which defines the cost of unmet need. Future work is needed to explore the use of AS while accounting for the dynamic nature of disease spread and resource availability.

@@

In conclusion, we argue that when possible, the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
the specific decision-making context. Defaulting to standard forecast evaluation metrics can mask the utility (or
disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
for which they are able to offer useful guidance. New collaborative work between decision-makers and modeling teams is
needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or
simulation exercises that could be used to inform further development of new or alternative scoring metrics. We see this
work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling
applied epidemiological forecasting with public health decision making.

@@

To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a Bayes act solving the
AP \eqref{eqn:AP-long} --- that is, we must find the specific resource allocations for each location that are determined by
the forecast $F$ under the resource constraint $K$. Assuming we have reliable means of calculating quantiles
$q_{\alpha,F_i}$ of the marginal forecasts $F_i$, these allocations are given by $q_{\tau^{\star},F_i}$ where
$\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}. This equation is not analytically tractable in general, but because the left-hand side of \eqref{eqn:quantiles-sum-to-K} is non-decreasing in $\tau$, it is straightforward to find arbitrarily accurate approximations to a solution using an iterative bisection method.

We have
implemented such a method along with the resulting score computations in the R package \verb`alloscore`
\citep{gerding-alloscore} which provided all allocation score values used in the analysis of section
\ref{sec:application}.
The procedure begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as
$[0,\max_{i}F_i(K)]$) that clearly contains the solution $\tau^{\star}$. At each step $j$ of the algorithm, we evaluate
the total allocation $\sum_{i=1}^N q_{\tau_{M,j},F_i}$ at the midpoint of the search interval,
$\tau_{M,j} = \frac{1}{2}(\tau_{L,j} + \tau_{U,j})$ and continue the search on the narrowed sub-interval
\begin{align}
[\tau_{L,j+1},\tau_{U,j+1}] =
\begin{cases}
[\tau_{L,j}, \tau_{M,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} \geq K \\
[\tau_{M,j}, \tau_{U,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} < K.
\end{cases} \nonumber
\end{align}
This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ 
for a suitably small $\varepsilon>0$.

Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref {sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f. \eqref{eqn:exp-quantile}).

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
\includegraphics[width=\maxwidth]{figure/bisection-1} \caption{Illustrations of bisection method for the examples of Section \ref{sec:methods.overview}. In each plot the thinner black lines are the quantile functions of the exponential forecasts for Location 1, with $\sigma_1=1$ (lower), and Location 2, with $\sigma_2=4$ (upper), while the thicker black line gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right. The red segments show the iterates $\tau_{M,j}$ and their associated total resource demands as given by the function \texttt{allocate} from the package \texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$. (The $\tau_{M,j}$ are in this case just sums of terms $\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f. Figure \ref{fig:exp_alloc_example} ).}\label{fig:bisection}
\end{figure}
